### 2022-08-03 software-table.tex
very minor changes

\toprule
\multicolumn{2}{l}{Software infrastructure}\\
\midrule
\texttt{demes-python} &
    A Python library for loading, saving, and working with
    Demes models. Includes support for converting to and from
    \texttt{ms}~\citep{hudson2002generating}
    (\url{https://github.com/popsim-consortium/demes-python}).\\

\texttt{demes-c} &
    A C library for parsing Demes YAML descriptions
    (\url{https://github.com/grahamgower/demes-c}). \\

\texttt{demes-rust} &
    A Demes parser in Rust
    (\url{https://github.com/molpopgen/demes-rs}). \\

\texttt{demes-julia} &
    A parser in Julia
    (\url{https://github.com/apragsdale/Demes.jl}). \\

\texttt{demesdraw} &
    A Python library for visualizing Demes models
    (\url{https://github.com/grahamgower/demesdraw}). \\

\midrule
\multicolumn{2}{l}{Methods using Demes as input/output format}\\
\midrule

% TOOL AUTHORS: please add a row to this table to summarise what the software
% does, and how it implements demes.

\dadi &
   Optimizes parameters in models of demographic history and distributions of fitness effects using SFS~\citep{gutenkunst2009inferring}. Can simulate SFS from Demes models.
    \\

\demesslim &
    Loads Demes models into the \slim\ forward simulator \citep{haller2019slim}.\\

\fwdpy &
   Simulates the Wright-Fisher model forward in time \citep{thornton2014cpp,thornton2019-nu}. Demes is the preferred format for specifying a demographic model.\\

\gadma &
    Infers models of demographic history \citep{noskova2020gadma}. Outputs Demes models and visualizations. \\

\moments &
    Optimizes parameters in models of demographic history using SFS and linkage disequilibrium statistics
    \citep{jouganous2017inferring,ragsdale2019models}. Models to be optimized can be specified in Demes.\\

\msprime &
    Simulates population genetic models using tree sequences~\citep{kelleher2016efficient,kelleher2020coalescent,baumdicker2021-iu}. Demographic history models can be specified using Demes.\\

\bottomrule